---
title: "pre_process_human"
author: "Helen_Liu"
date: "13/07/2022"
output: html_document
---

```{r load packages}
library(tidyverse)
library(ggplot2)
library(bruceR)
library(ggpubr)
#
source('convenient_functions.R')
source('Function_for_violin_plot.R')
getwd()
```


```{r load data}
df <- read.csv("~/Desktop/nma2022_project_RL_riskychoice/data/human_data.csv")
#str(df)
```


```{r load data}
#see how many human participants we have
length(unique(df$Subject)) #28
```


```{r preprocessing}
#check data
#xxx <- df %>% filter(Subject ==1)
#xxx2 <- xxx %>% filter(TrialType != "NULL")
#unique(xxx2$Block)
#unique(xxx$Trial) #48 trials in each blocks
#length(xxx2$Trial)
#length(unique(xxx2$StimType))

###
df_new <- df %>% dplyr::select(
  "participant_id" = "Subject",
  "TrialType",
  #three types of trials
  "StimType",
  #stimulus
  "Block",
  #6 blocks
  "trial_index" = "Trial",
  #trial number
  "response" = "Choice",
  #participants' choice
  "RewardAmount",
  #amount of reward participants gets
  "CumulativeReward" #the cumulated reward participants gets
) %>%
  dplyr::filter(TrialType != "NULL") %>% #exclude the breaks
  dplyr::mutate(
    n_Block = case_when(
      #correct the block numbers
      Block == "1" ~ "1",
      Block == "3" ~ "2",
      Block == "5" ~ "3",
      Block == "7" ~ "4",
      Block == "9" ~ "5",
      Block == "11" ~ "6"
    )
  )  %>%
  dplyr::mutate(
    TypeTrial = case_when(
      #mark types of trials
      TrialType == "" ~ "forced_choice",
      # only one door
      TrialType == "Decision" ~ "Decision Trial",
      #chose between the safe and risky options with equal expected values 12/34
      TrialType == "Catch" ~ "Catch Trial" # chose between options of unequal expected value (high versus low) 13/14/23/24
    )
  ) %>%
  dplyr::filter(TypeTrial != "forced_choice") %>% #exclude one door condition
  dplyr::mutate(
    expected_response = case_when(
      #mark expected response (maximize reward)
      StimType == "102" | StimType == "201" ~ "Door2",
      #decision
      StimType == "103" | StimType == "301" ~ "Door3",
      #catch
      StimType == "104" | StimType == "401" ~ "Door4",
      #catch
      StimType == "203" | StimType == "302" ~ "Door3",
      #catch
      StimType == "204" | StimType == "402" ~ "Door4",
      #catch
      StimType == "304" | StimType == "403" ~ "Door4" #decision
    )
  )  %>%
  dplyr::mutate(is_correct =
                  case_when(
                    #mark accuracy
                    response == expected_response ~ "1",
                    response != expected_response ~ "0"
                  )) %>%
  dplyr::select(-Block, -TrialType)
head(df_new)
```


```{r decision trials}
df_decision <- df_new %>%
  dplyr::filter(TypeTrial == "Decision Trial") %>% #only select decision
  dplyr::mutate(
    ValueCondition =
      case_when(
        #mark high vs low value
        StimType == "102" ~ "Low value",
        StimType == "201" ~ "Low value",
        StimType == "304" ~ "High value",
        StimType == "403" ~ "High value",
      )
  ) %>%
  dplyr::select(-RewardAmount, -CumulativeReward)
head(df_decision)
#length(unique(df_new$participant_id)) #check participants’ number
```


```{r plot descriptive in  decision}
colnames(df_decision)
stats <- df_decision %>%
  dplyr::group_by(participant_id, n_Block,ValueCondition) %>%
  mutate_at("is_correct", as.numeric) %>%
  dplyr::summarise(n = n(),
                   p_risky = mean(is_correct)) %>%
  mutate_at(vars(n_Block, participant_id,ValueCondition), as.factor)
#save stats
#save_data(stats,'df.csv','~/Desktop/')
```


```{r ploting}
# plot results
###save plot
getwd()

tiff(
  file = "descript_human.tiff",
  res = 800,
  width = 4000,
  height = 2500
)#save tiff
###
colnames(stats)
ggplot(
  data = stats,
  mapping = aes(
    x = n_Block,
    y = p_risky,
    group = participant_id,
    color = participant_id
  )
) +
  facet_wrap(~ ValueCondition) +
  theme(strip.background = element_blank(),
        strip.text.x = element_text(size = 12, color = "Black")) +
  geom_line(alpha = 0.15) +
  geom_point(alpha = 0.1, size = 1) +
  stat_summary(
    aes(y = p_risky, group = 1),
    fun = mean,
    colour = "black",
    size = 0.6,
    geom = "line",
    alpha = 0.7,
    group = 1
  ) +
  stat_summary(
    aes(y = p_risky, group = 1),
    fun.data = "mean_se",
    geom = "pointrange",
    colour = "black",
    size = 0.2,
    group = 1
  ) + # adding error bars (standard error)
  labs(x = "Block", y = "Poportion of risky choice") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position = "none",
    axis.line = element_line(colour = "black")
  )
#save graph
#dev.off()
```

#violin plot

```{r violin plot}
Data_summary <- summarySE(stats, measurevar="p_risky", groupvars=c("ValueCondition","n_Block"))
Data_summary$n_Block <- as.numeric(Data_summary$n_Block)
Data_summary$n_Block_j <- jitter(Data_summary$n_Block, amount = 0.1)

legend_title <- "Value condition"
#
tiff(
  file = "descript_human.tiff",
  res = 800,
  width = 4000,
  height = 2500
)#save tiff
###
ggplot(data = stats,
       mapping = aes(
         x = as.factor(n_Block),
         y = p_risky,
         fill = ValueCondition
       )) +
  geom_split_violin(trim = T,
                    color = "white",
                    scale = "area") + #绘制分半的小提琴图
  geom_point(
    data = Data_summary,
    aes(x = as.factor(n_Block), y = p_risky),
    pch = 19,
    position = position_dodge(0.5),
    size = 1
  ) +
  geom_errorbar(
    data = Data_summary,
    aes(ymin = p_risky - ci, ymax = p_risky + ci),
    width = 0.05,
    position = position_dodge(0.5),
    color = "black",
    alpha = 0.8,
    size = 0.5
  ) + #add error bar  #绘制均值为点图
  geom_line(
    data = Data_summary %>% filter(ValueCondition == "High value"),
    mapping = aes(
      x = as.numeric(n_Block) - 0.11,
      y = p_risky
    ),
    #linetype = "dashed",
    alpha = 0.4,
    show.legend = F
  ) +
  geom_line(
    data = Data_summary %>% filter(ValueCondition == "Low value"),
    mapping = aes(
      x = as.numeric(n_Block) + 0.11,
      y = p_risky,
    ),
    linetype = "dashed",
    alpha = 0.4,
    show.legend = F
  ) +
  scale_fill_manual(legend_title,
                    values = c("#56B4E9", "#E69F00")) +
  labs(
    x = ("Blocks"),
    y = ("Poportion of risky choice"),
    x = NULL,
    #title = "Poportion of risky choice under different expected value condition"
  ) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black")
  ) +
  stat_compare_means(
    aes(group = ValueCondition),
    label = "p.signif",
    method = "t.test",
    paired = TRUE,
    label.y = max(stats$p_risky),
    hide.ns = T
  ) # add paired t-test results
#dev.off()
```

```{r ploting cummulated rewards}
df_cu <- df_new %>%
  dplyr::select("participant_id",
                "RewardAmount",
                "TypeTrial",
                "n_Block") %>%
  dplyr::group_by(participant_id,
                  n_Block) %>%
  dplyr::mutate(Cumulative = cumsum(RewardAmount)) %>% #add cumulative rewards)
  dplyr::mutate(trials = row_number()) %>% # add number of trials
  dplyr::ungroup()%>% 
  dplyr::rename("Block" = "n_Block")%>% 
  mutate_at("participant_id", as.factor) %>%
  mutate_at(vars(Cumulative, RewardAmount), as.numeric)


tiff(
  file = "descript2_human.tiff",
  res = 800,
  width = 4500,
  height = 3500
  
)#save tiff
ggplot(
  data = df_cu,
  mapping = aes(
    x = trials,
    y = Cumulative,
    group = participant_id,
    color = participant_id
  ))+
  geom_line(alpha = 0.15)+
  scale_y_continuous(expand = c(0,0))+
  facet_grid(~Block, labeller = label_both)+
  stat_summary(
    aes(y = Cumulative, group = 1),
    fun = mean,
    colour = "black",
    size = 0.6,
    geom = "line",
    alpha = 0.7,
    group = 1
  ) +  
  labs(x = "Trials", y = "Cumulative rewards") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position = "none",
    axis.line = element_line(colour = "black")
  )
 dev.off()
```


```{r ploting color options}
# individual
df_color <- df_new %>%
  dplyr::mutate(
    Response = case_when(
      response == "Door1" ~ "1",
      response == "Door2" ~ "2",
      response == "Door3" ~ "3",
      response == "Door4" ~ "4"
    )
  ) %>%
  select(-expected_response,-is_correct,-response) %>%
  group_by(participant_id, n_Block, Response) %>%
  dplyr::summarise(n = n()) %>%
  mutate(p = n / base::sum(n)) %>%
  mutate_at(vars(n_Block, participant_id), as.factor) %>%
  mutate_at("p", as.numeric)

#colnames(df_color)
df_color
#
tiff(
  file = "ppc_human.tiff",
  res = 800,
  width = 10000,
  height = 3500)
#
ggplot(data = df_color,
       aes(x = n_Block, y = 100 * p, fill = Response)) + #geom_col和geom_bar这两条命令都可以绘制堆叠柱形图
  #geom_col(position = 'stack', width = 0.6)+
  geom_bar(position = "stack",
           stat = "identity",
           width = 0.6) +
  facet_grid( ~ participant_id) +
  theme(strip.text.x = element_blank()) + #remove Labels from Facet Plot
  scale_fill_manual(values = c(
    "1" = "#F0BE26", #yellow
    "2" = "#296613",#green
    "3" = "#CC1F19" ,#red
    "4" = "#1B3D80" #blue
  )) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Blocks",
       y = "Relative proportion",
       fill = "Door options",
       title = " Relative proportion of door choices across blocks (individual level)")
 dev.off()
```


```{r ploting color options}
#average 
df_color2 <- df_new %>%
  dplyr::mutate(
    Response = case_when(
      response == "Door1" ~ "1",
      response == "Door2" ~ "2",
      response == "Door3" ~ "3",
      response == "Door4" ~ "4"
    )
  ) %>%
  select(-expected_response, -is_correct, -response) %>%
  group_by(n_Block,Response) %>%
  dplyr::summarise(n = n()) %>%
  mutate(p = n / base::sum(n))%>%
  mutate_at(vars(n_Block),as.factor)%>%
  mutate_at("p",as.numeric)

ggplot(data = df_color2,
       aes(x = n_Block, y = 100 * p, fill = Response)) + #geom_col和geom_bar这两条命令都可以绘制堆叠柱形图
  #geom_col(position = 'stack', width = 0.6)+
  geom_bar(position = "stack",
           stat = "identity",
           width = 0.6) +
  scale_fill_manual(values = c(
    "1" = "#F0BE26",
    "2" = "#296613",
    "3" = "#CC1F19" ,
    "4" = "#1B3D80"
  )) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Blocks",
       y = "Relative proportion",
       fill = "Door options",
       title = " Relative proportion of door choices across blocks (average)")
```

# Group-level Relative proportion of door choices across blocks (in decision trials)

```{r ploting color options}
#only decision
df_color3 <- df_new %>%
  dplyr::mutate(
    Response = case_when(
      response == "Door1" ~ "1",
      response == "Door2" ~ "2",
      response == "Door3" ~ "3",
      response == "Door4" ~ "4"
    )
  ) %>%
  filter(TypeTrial == "Decision Trial") %>%
  select(-expected_response, -is_correct, -response) %>%
  group_by(n_Block,Response) %>%
  dplyr::summarise(n = n()) %>%
  mutate(p = n / base::sum(n))%>%
  mutate_at(vars(n_Block),as.factor)%>%
  mutate_at("p",as.numeric)

ggplot(data = df_color3,
       aes(x = n_Block, y = 100 * p, fill = Response)) + 
  geom_bar(position = "stack",
           stat = "identity",
           width = 0.6) +
  #facet_grid( ~ participant_id) +
  #theme(strip.text.x = element_blank()) + #remove Labels from Facet Plot
  scale_fill_manual(values = c(
    "1" = "#F0BE26",
    "2" = "#296613",
    "3" = "#CC1F19" ,
    "4" = "#1B3D80"
  )) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Blocks",
       y = "Relative proportion",
       fill = "Door options",
       title = " Relative proportion of door choices across blocks (in decision trials)")
```
```{r}
#only decision
df_color4 <- df_new %>%
  dplyr::mutate(
    Response = case_when(
      response == "Door1" ~ "1",
      response == "Door2" ~ "2",
      response == "Door3" ~ "3",
      response == "Door4" ~ "4"
    )
  ) %>%
  filter(TypeTrial == "Decision Trial") %>%
  select(-expected_response, -is_correct, -response) %>%
  group_by(participant_id,n_Block,Response) %>%
  dplyr::summarise(n = n()) %>%
  mutate(p = n / base::sum(n))%>%
  mutate_at(vars(n_Block),as.factor)%>%
  mutate_at("p",as.numeric)

ggplot(data = df_color4,
       aes(x = n_Block, y = 100 * p, fill = Response)) + #geom_col和geom_bar这两条命令都可以绘制堆叠柱形图
  #geom_col(position = 'stack', width = 0.6)+
  geom_bar(position = "stack",
           stat = "identity",
           width = 0.6) +
  facet_grid( ~ participant_id) +
  theme(strip.text.x = element_blank()) + #remove Labels from Facet Plot
  scale_fill_manual(values = c(
    "1" = "#F0BE26",
    "2" = "#296613",
    "3" = "#CC1F19" ,
    "4" = "#1B3D80"
  )) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Blocks",
       y = "Relative proportion",
       fill = "Door options",
       title = " Relative proportion of door choices across blocks in decison trials (individual level)")
```

#calculate correlation 

```{r}
df_cor.wide <- df_color4 %>% dplyr::filter(Response == 2|Response == 3) %>%
  mutate_at(vars("participant_id","Response"),as.factor) %>%
  group_by(participant_id,Response) %>%
  summarise(mean_p = mean(p))%>%
  ungroup() %>% pivot_wider(id_cols = "participant_id", names_from = "Response", values_from = "mean_p")
  
## calculate corr
df_cor.wide[-1] %>% rstatix::cor_test("2":"3")
#no correlation  
```
  

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

