---
title: "pre_process_human"
author: "Helen_Liu"
date: "13/07/2022"
output: html_document
---

```{r load packages}
library(tidyverse)
library(ggplot2)
library(bruceR)
#
source('convenient_functions.R')
getwd()
```


```{r load data}
df <- read.csv("~/Desktop/nma2022_project_RL_riskychoice/data/human_data.csv")
#str(df)
```


```{r load data}
#see how many human participants we have
length(unique(df$Subject)) #28
```


```{r preprocessing}
#check data
#xxx <- df %>% filter(Subject ==1)
#xxx2 <- xxx %>% filter(TrialType != "NULL")
#unique(xxx2$Block)
#unique(xxx$Trial) #48 trials in each blocks
#length(xxx2$Trial)
#length(unique(xxx2$StimType))

###
df_new <- df %>% dplyr::select(
  "participant_id" = "Subject",
  "TrialType",
  #three types of trials
  "StimType",
  #stimulus
  "Block",
  #6 blocks
  "trial_index" = "Trial",
  #trial number
  "response" = "Choice",
  #participants' choice
  "RewardAmount",
  #amount of reward participants gets
  "CumulativeReward" #the cumulated reward participants gets
) %>%
  dplyr::filter(TrialType != "NULL") %>% #exclude the breaks
  dplyr::mutate(
    n_Block = case_when(
      #correct the block numbers
      Block == "1" ~ "1",
      Block == "3" ~ "2",
      Block == "5" ~ "3",
      Block == "7" ~ "4",
      Block == "9" ~ "5",
      Block == "11" ~ "6"
    )
  )  %>%
  dplyr::mutate(
    TypeTrial = case_when(
      #mark types of trials
      TrialType == "" ~ "forced_choice",
      # only one door
      TrialType == "Decision" ~ "Decision Trial",
      #chose between the safe and risky options with equal expected values 12/34
      TrialType == "Catch" ~ "Catch Trial" # chose between options of unequal expected value (high versus low) 13/14/23/24
    )
  ) %>%
  dplyr::filter(TypeTrial != "forced_choice") %>% #exclude one door condition
  dplyr::mutate(
    expected_response = case_when(
      #mark expected response (maximize reward)
      StimType == "102" | StimType == "201" ~ "Door2",
      #decision
      StimType == "103" | StimType == "301" ~ "Door3",
      #catch
      StimType == "104" | StimType == "401" ~ "Door4",
      #catch
      StimType == "203" | StimType == "302" ~ "Door3",
      #catch
      StimType == "204" | StimType == "402" ~ "Door4",
      #catch
      StimType == "304" | StimType == "403" ~ "Door4" #decision
    )
  )  %>%
  dplyr::mutate(is_correct =
                  case_when(
                    #mark accuracy
                    response == expected_response ~ "1",
                    response != expected_response ~ "0"
                  )) %>%
  #dplyr::filter(StimType != 103&301) %>% #20vs60
  dplyr::select(-Block, -TrialType)
head(df_new)
#length(unique(df_new$participant_id)) #check participants’ number
```


```{r descriptive}
colnames(df_new)
stats <- df_new %>%
  dplyr::group_by(participant_id, n_Block, TypeTrial) %>%
  mutate(across("is_correct", as.numeric)) %>%
  dplyr::summarise(n = n(),
                   acc = mean(is_correct)) %>%
  mutate_at(vars(n_Block, participant_id), as.factor)
length(unique(stats$participant_id))

#save stats
#save_data(stats,'df.csv','~/Desktop/')
```


```{r ploting}

# plot results
###save plot
getwd()

tiff(
  file = "descript_human.tiff",
  res = 800,
  width = 4000,
  height = 2500
)#save tiff
###
colnames(stats)
ggplot(
  data = stats,
  mapping = aes(
    x = n_Block,
    y = acc,
    group = participant_id,
    color = participant_id
  )
) +
  facet_wrap(~ TypeTrial) +
  theme(strip.background = element_blank(),
        strip.text.x = element_text(size = 12, color = "Black")) +
  geom_line(alpha = 0.15) +
  geom_point(alpha = 0.1, size = 1) +
  stat_summary(
    aes(y = acc, group = 1),
    fun = mean,
    colour = "black",
    size = 0.6,
    geom = "line",
    alpha = 0.7,
    group = 1
  ) +
  stat_summary(
    aes(y = acc, group = 1),
    fun.data = "mean_se",
    geom = "pointrange",
    colour = "black",
    size = 0.2,
    group = 1
  ) + # adding error bars (standard error)
  labs(x = "Block", y = "Poportion of risky choice") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position = "none",
    axis.line = element_line(colour = "black")
  )
#save graph
dev.off()
#
p

```


```{r ploting}

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

